DeepSphere: a graph-based spherical CNN

### 2019-11-11 rebuttal.md

As a non-expert of spherical CNN, I don't understand clearly the gap between the result Theorem 3.1 and showing the rotation equivalence. It would be nice to add some counterexample (i.e., in what situation the proposed approach does not have rotational equivalence while Theorem 3.1 holds).

### Answer

We thank the reviewer for their time assessing our work and their constructive feedback.

We prepared a revised manuscript, to be uploaded shortly, containing a deeper theoretical discussion closing the gap between the result Theorem 3.1 and rotation equivariance. In a small proposition, to be added after theorem 3.1, we precise mathematically the relationship between these two concepts. In short, we proved that, if theorem 3.1 holds, our graph Laplacian L commutes with any rotation operator R in the limit of infinite sampling (pointwise), i.e. LRf(x)-RLf(x) \to 0, thus answering the reviewer's concerns about this subject.

### 2019-11-11 rebuttal.md

Minor:
- A figure detailing the parameters and setup for theorem 3.1 and figure 2 would be useful.
- The statement on the dispersion of the sampling sequence states “the smallest ball in \R^3 containing \sigma_i”, but I believe it should be “containing only \sigma_i”.

### Answer

We thank the reviewer for their time assessing our work and their constructive feedback.

We deliberately excluded experiments on omnidirectional imagery. In our opinion, those don't possess full spherical symmetries as gravity is orienting the objects. We encourage the reviewer to check the work of Khasanova and Frossard, who explicitly designed graph-based spherical CNNs for omnidirectional imaging. In [1], they designed a graph that yields an equivariant convolution to a subgroup of SO(3). Longitudinal rotations are equivariant by construction of the equiangular sampling, and they optimized the graph for latitudinal equivariance. Their scheme is presented in section 3.2 of our paper. While their convolution is not equivariant to the whole of SO(3), that is not an issue for this application as gravity prevents objects from rotating around the third axis. It may even be beneficial. Moreover, the orientation given by gravity allows to factorize the spherical graph and design anisotropic filters [2].

Radius or kNN graphs are means to get a sparse graph for O(n) matrix multiplication, instead of O(n²) for the full distance-based similarity graph. We believe that the choice of one or the other doesn't really matter. Sparsification can be seen as a numerical approximation that replaces small values by zeros. The kNN scheme is often preferred in practice as the choice of k is directly linked to the computational cost, while the choice of a radius large enough to avoid disconnected vertices might include many more edges than necessary on denser areas.

Thanks for pointing out an unclear statement about the dispersion of the sampling sequence. d_i should be understood as the largest distance between the center x_i and any point on the surface sigma_i. Hence, we define d_i to be the radius of the smallest ball centered at x_i containing sigma_i. We'll clarify.

@@

As a non-expert of spherical CNN, I don't understand clearly the gap between the result Theorem 3.1 and showing the rotation equivalence. It would be nice to add some counterexample (i.e., in what situation the proposed approach does not have rotational equivalence while Theorem 3.1 holds).

### Answer

We thank the reviewer for their time assessing our work and their constructive feedback.

We prepared a revised manuscript, to be uploaded shortly, containing a deeper theoretical discussion closing the gap between Theorem 3.1 and rotation equivariance. In a small proposition, to be added after theorem 3.1, we precise mathematically the relationship between these two concepts. In short, we proved that, if theorem 3.1 holds, our graph Laplacian L commutes with any rotation operator R in the limit of infinite sampling (pointwise), i.e., |LRf(x) - RLf(x)| -> 0, thus answering the reviewer's concerns about this subject.

### 2019-11-11 main.tex
last pass on the rebuttal

\paragraph{Theoretical convergence.}
We extend the work of Belkin and Nyiogi to a sufficiently regular, deterministic sampling scheme. \mart{Here we insert the definition of the extended graph laplacian and of the rotation operator Rg} Following their setting, we define the extended graph Laplacian operator as a linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\begin{equation}
	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^tf(y) := \frac{1}{n}\sum_{i=0}^{n-1} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
Note that we have the following relation $\b{L}_n^t \b{f} = T_\V L_n^t f$.
Given a sampling scheme $\V = \{x_0, \dots, x_{n-1}\}$, we also define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area and $d_i$ the radius of the smallest ball in $\R^3$ containing $\sigma_i$ and centered at $x_i$. Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$. Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere 
\begin{equation}\label{eq:rotation operator}
\mathcal R_g f(x) := f(g^{-1}(x)).
\end{equation}
\begin{theorem}
	For a sampling $\V = \{x_i\in\S^2\}_{i=0}^{n-1}$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that 
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$  
	\begin{equation*}
		\lim_{n\to\infty}\frac{1}{4\pi t_n^2} L_n^{t_n}f(y) = \frac{1}{|\S^2|}\triangle_{\S^2}f(y)
	\end{equation*} 
% 	for $n\to\infty$.
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, $\b{L}$ converges pointwise towards the Laplace-Beltrami operator ($\Delta_{\S^2}$) in the sense explained in \cite{belkin2007convergence}. \mart{here we insert the link between consistency and equivariance} This operator commutes with rotation and hence this convergence is a strong hint for equivariance, as formalized in the following proposition.
%%%%%%%%%%%%%%%%%%%%%%%  link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{prop}\label{prop:equivariance}
Under the hypothesis of theorem \ref{theo:pointwise convergence for a regular sampling}, we have the following pointwise convergence
\begin{equation*}
\forall x\in\mathbb S^2\quad\left| \mathcal R_g L_n^{t_n} f (x) - L_n^{t_n} \mathcal R_g f(x) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{prop}
%%%%%%%%%%%%%%%%%%%%%%% end of link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\paragraph{Theoretical convergence.}
We extend the work of Belkin and Nyiogi to a sufficiently regular, deterministic sampling scheme. \mart{Here we insert the definition of the extended graph laplacian and of the rotation operator Rg} Following their setting, we define the extended graph Laplacian operator as a linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\begin{equation}
	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^tf(y) := \frac{1}{n}\sum_{i=0}^{n-1} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
Note that we have the following relation $\b{L}_n^t \b{f} = T_\V L_n^t f$.
Given a sampling scheme $\V = \{x_0, \dots, x_{n-1}\}$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ the radius of the smallest ball in $\R^3$ containing $\sigma_i$ and centered at $x_i$.%\footnote{$d_i$ should be understood as the largest distance between the center $x_i$ and any point on the surface $sigma_i$.}
Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.
Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere
\begin{equation}\label{eq:rotation operator}
	\mathcal R_g f(x) := f(g^{-1}(x)).
\end{equation}
\begin{theorem}
	For a sampling $\V = \{x_i\in\S^2\}_{i=0}^{n-1}$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that 
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$  
	\begin{equation*}
		\lim_{n\to\infty}\frac{1}{4\pi t_n^2} L_n^{t_n}f(y) = \frac{1}{|\S^2|}\triangle_{\S^2}f(y).
	\end{equation*} 
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, $\b{L}$ converges pointwise towards the Laplace-Beltrami operator ($\Delta_{\S^2}$) in the sense explained in \cite{belkin2007convergence}. \mart{here we insert the link between consistency and equivariance} This operator commutes with rotation and hence this convergence is a strong hint for equivariance, as formalized in the following proposition.
%%%%%%%%%%%%%%%%%%%%%%%  link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{prop}\label{prop:equivariance}
Under the hypothesis of theorem \ref{theo:pointwise convergence for a regular sampling}, we have the following pointwise convergence
\begin{equation*}
\forall x\in\mathbb S^2\quad\left| \mathcal R_g L_n^{t_n} f (x) - L_n^{t_n} \mathcal R_g f(x) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{prop}
%%%%%%%%%%%%%%%%%%%%%%% end of link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%
